TKT RNAseq Experiment: Differential Expression Analysis

### 2025-04-11 GeneralDataFiles/CompleteMetabolicNetwork.Rmd
Updated Russo Data

G85R.geneFC = read.csv(paste0(git.dir, "TKT.EdgeR.FCTable.csv"), row.names = 1)
G85R.metabFC = read.csv(paste0(git.dir, "MetaboliteFCs.csv"), row.names = 1)

@@

temp = colorRampPalette(c('royalblue',"dodgerblue",'red', "firebrick"))(401)

@@

node.colors = G85R.metabFC[nodes$data.id, "GR.CxOE"]
node.colors = temp[(200*node.colors/max(abs(na.exclude(node.colors)))) + 201]
node.colors[na.action(na.exclude(node.colors))] = "white"

@@

visNetwork(nodes, edges)%>%
  visNodes(physics = F)%>%
  visLayout(hierarchical=F, improvedLayout=T)%>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE)%>%
  visEdges(arrows = visedges$arrows)

>>>>

GeneIDKey = read.csv(paste0(git.dir, "GeneIDKey.csv"), row.names = 1)
G85R.geneFC = read.csv(paste0(git.dir, "TKT.EdgeR.FCTable.csv"), row.names = 1)
G85R.metabFC = read.csv(paste0(git.dir, "MetaboliteFCs.csv"), row.names = 1)
G85R.metabFC[,2:ncol(G85R.metabFC)] = log2(G85R.metabFC[,2:ncol(G85R.metabFC)])

@@

temp = colorRampPalette(c("dodgerblue", "grey",'red'))(401)

```{r echo = F}
##"GRxWT.C"  "GRxWT.Df" "GRxWT.OE" "GR.CxDf" "WT.CxDf"  "GR.CxOE"  "WT.CxOE"
metab = "GR.CxDf"

enzyme = "GRF.CxDf"

nodes[nodes$shape == "dot", "label"] = NA

node.colors = (G85R.metabFC[nodes$data.id, metab])
node.colors = temp[(200*node.colors/max(abs(na.exclude(node.colors)))) + 201]
node.colors[na.action(na.exclude(node.colors))] = "white"

nodes$title = paste0(nodes$id,
                     "<br>FC: ", 
                     signif((G85R.metabFC[nodes$data.id, metab]), 3))

edges$title = paste0(GeneIDKey[edges$FBgn, "Symbol"],
                     "<br>FC: ", 
                     signif(G85R.geneFC[edges$FBgn,
                                 enzyme],3))

edges.colors = G85R.geneFC[edges$FBgn, enzyme]
edges$width = abs(edges.colors) * 10

edges.colors = temp[(200*edges.colors/max(abs(na.exclude(edges.colors)))) + 202]
edges.colors[na.action(na.exclude(edges.colors))] = "white"

edges$color = edges.colors

#edges = edges[c(224,225, 226,227),]
#edges$title = c("A","B","C", "D")

visNetwork(nodes, edges)%>%
  visNodes(physics = F)%>%
  #visLayout(hierarchical=F, improvedLayout=T)%>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = T)%>%
  visEdges(arrows = visedges$arrows)

@@

nodes[nodes$shape == "circle", "size"] = 20
nodes$shape = "dot"

visNetwork(nodes, edges)%>%
  visNodes(physics = F)

@@

### 2025-04-11 TKT_RNAseq/R Code/EdgeR_Analysis.Rmd
update
Updated Russo Data

##Clean Trait Data and calculate CPMs
countdata=read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT.counttable.csv",row.names=1)
groups=read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/All.TKT.metadata.csv",row.names=1)

convert=read.csv("https://raw.githubusercontent.com/DavidRandLab/Santiago-et-al-2021-BMC-Genomics/main/Data%20Files/FBgnConversionTable.csv",row.names=1)

##normalize data
countdata=countdata[, row.names(groups)]
x <- countdata
gr <- factor(groups$Group)
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
cpmdata=cpm(z)

@@

##GR.F TKT control vs TKT altered
compare = makeContrasts(TktDfGR.F-GR.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.F.CxDf=DEout
sGR.F.CxDf=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.F-GR.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.F.CxOE=DEout
sGR.F.CxOE=DEout[DEout$FDR<.05,]

##WT.F TKT control vs TKT altered
compare = makeContrasts(TktDfWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.F.CxDf=DEout
sWT.F.CxDf=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.F.CxOE=DEout
sWT.F.CxOE=DEout[DEout$FDR<.05,]

##WT.F vs GR.F
compare = makeContrasts(GR.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.F=DEout
sGRxWT.F=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.F-TktDfWT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.FDf=DEout
sGRxWT.FDf=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.F-TktOEWT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.FOE=DEout
sGRxWT.FOE=DEout[DEout$FDR<.05,]

##GR.M TKT control vs TKT altered
compare = makeContrasts(TktDfGR.M-GR.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.M.CxDf=DEout
sGR.M.CxDf=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.M-GR.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.M.CxOE=DEout
sGR.M.CxOE=DEout[DEout$FDR<.05,]

##WT.M TKT control vs TKT altered
compare = makeContrasts(TktDfWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.M.CxDf=DEout
sWT.M.CxDf=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.M.CxOE=DEout
sWT.M.CxOE=DEout[DEout$FDR<.05,]

##WT.M vs GR.M
compare = makeContrasts(GR.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.M=DEout
sGRxWT.M=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.M-TktDfWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.MDf=DEout
sGRxWT.MDf=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.M-TktOEWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.MOE=DEout
sGRxWT.MOE=DEout[DEout$FDR<.05,]

TKT.EdgeR = data.frame(GRF.CxDf = GR.F.CxDf[row.names(cpmdata),'FDR'],
                       GRF.CxOE = GR.F.CxOE[row.names(cpmdata),'FDR'],

                       WTF.CxDf = WT.F.CxDf[row.names(cpmdata),'FDR'],
                       WTF.CxOE = WT.F.CxOE[row.names(cpmdata),'FDR'],

                       GRxWT.FC = GRxWT.F[row.names(cpmdata),'FDR'],
                       GRxWT.FDf = GRxWT.FDf[row.names(cpmdata),'FDR'],
                       GRxWT.FOE = GRxWT.FOE[row.names(cpmdata),'FDR'],

                       GRM.CxDf = GR.M.CxDf[row.names(cpmdata),'FDR'],
                       GRM.CxOE = GR.M.CxOE[row.names(cpmdata),'FDR'],

                       WTM.CxDf = WT.M.CxDf[row.names(cpmdata),'FDR'],
                       WTM.CxOE = WT.M.CxOE[row.names(cpmdata),'FDR'],

                       GRxWT.MC = GRxWT.M[row.names(cpmdata),'FDR'],
                       GRxWT.MDf = GRxWT.MDf[row.names(cpmdata),'FDR'],
                       GRxWT.MOE = GRxWT.MOE[row.names(cpmdata),'FDR'])

row.names(TKT.EdgeR) = row.names(cpmdata)
##write.csv(TKT.EdgeR, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT.EdgeR.Table.csv")

MDSdata = plotMDS(z, top = 13582, col=as.numeric(factor(gr)))

temp=convert[,"Symbol"]
temp2=setdiff(row.names(countdata),convert[,2])
names(temp)=convert[,2]
names(temp2)=temp2
convert=c(temp,temp2)

Comparisons = list(row.names(sGR.F.CxDf), row.names(sGR.F.CxOE),
                row.names(sWT.F.CxDf), row.names(sWT.F.CxOE),
                row.names(sGRxWT.F), row.names(sGRxWT.FDf), row.names(sGRxWT.FOE),
                row.names(sGR.M.CxDf), row.names(sGR.M.CxOE),
                row.names(sWT.M.CxDf), row.names(sWT.M.CxOE),
                row.names(sGRxWT.M), row.names(sGRxWT.MDf), row.names(sGRxWT.MOE))

GroupsToDEGs = matrix("", nrow = 28, ncol = 3)

GroupsToDEGs[,1] = c("GR.F", "TktDfGR.F",
                     "GR.F", "TktOEGR.F",
                     "WT.F", "TktDfWT.F",
                     "WT.F", "TktOEWT.F",
                     "GR.F", "WT.F",
                     "TktDfGR.F", "TktDfWT.F",
                     "TktOEGR.F", "TktOEWT.F",
                     "GR.M", "TktDfGR.M",
                     "GR.M", "TktOEGR.M",
                     "WT.M", "TktDfWT.M",
                     "WT.M", "TktOEWT.M",
                     "GR.M", "WT.M",
                     "TktDfGR.M", "TktDfWT.M",
                     "TktOEGR.M", "TktOEWT.M")

GroupsToDEGs[,2] = c("TktDfGR.F", "GR.F",
                     "TktOEGR.F", "GR.F", 
                     "TktDfWT.F", "WT.F", 
                     "TktOEWT.F", "WT.F", 
                     "WT.F", "GR.F", 
                     "TktDfWT.F", "TktDfGR.F", 
                     "TktOEWT.F", "TktOEGR.F", 
                     "TktDfGR.M", "GR.M", 
                     "TktOEGR.M", "GR.M", 
                     "TktDfWT.M", "WT.M", 
                     "TktOEWT.M", "WT.M", 
                     "WT.M", "GR.M", 
                     "TktDfWT.M", "TktDfGR.M", 
                     "TktOEWT.M", "TktOEGR.M")

GroupsToDEGs[,3] = c(1,1,2,2,3,3,4,4,5,5,6,6,7,7,
                     8,8,9,9,10,10,11,11,12,12,13,13,14,14)

>>>>

GeneIDKey = read.csv("https://raw.githubusercontent.com/johncsantiago/WhartonLab/refs/heads/master/GeneralDataFiles/GeneIDKey.csv", row.names = 1)

#raw count data directory
cd.dir = "/Users/johncsantiago/Documents/GitHub/Datasets/GBB_Count_Tables/"

GBB.countdata = read.csv(paste0(cd.dir, list.files(cd.dir)[1]), 
                   sep = "\t", row.names = 1)
colnames(GBB.countdata) = substr(list.files(cd.dir)[1], 1, 7)

i=2
while(i <= length(list.files(cd.dir))){
  temp.cd = read.csv(paste0(cd.dir, list.files(cd.dir)[i]), 
                     sep = "\t", row.names = 1)
  colnames(temp.cd) = substr(list.files(cd.dir)[i], 1, 7)
  temp.cd[,1] = temp.cd[row.names(GBB.countdata),]
  GBB.countdata = cbind(GBB.countdata, temp.cd)
  i = i + 1
}

GBB.groups = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/GBB_Metadata.csv", row.names = 1)

##Clean Trait Data and calculate CPMs
countdata=GBB.countdata
groups=GBB.groups

groups = groups[groups$Experiment == "GBB-KO",]

##normalize data
countdata= countdata[,row.names(groups)]
x <- countdata
gr <- factor(groups$Group)
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
cpmdata=cpm(z)

@@

##GR.F TKT control vs TKT altered
#compare = makeContrasts(TktDfGR.F-GR.F, levels=design)
#lrt = glmQLFTest(fit,contrast=as.vector(compare))
#G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
#DEout=G_X_E$table
#GR.F.CxDf=DEout
#sGR.F.CxDf=DEout[DEout$FDR<.05,]

MDSdata = plotMDS(z, 
                  top = 15000, 
                  col = as.numeric(factor(gr)), 
                  labels =  paste(groups$Group, groups$Replicate))

```{r echo = F}

countdata=GBB.countdata
groups=GBB.groups

groups = groups[groups$Experiment == "GBB-OE",]

##normalize data
countdata= countdata[,row.names(groups)]
x <- countdata
gr <- factor(groups$Group)
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
cpmdata=cpm(z)

fit <- glmQLFit(z, design)

MDSdata = plotMDS(z, 
                  top = 15000, 
                  col = as.numeric(factor(gr)), 
                  labels =  paste(groups$Group, groups$Replicate))
